### 2024-02-09 outputs/paper/paper.qmd
cleaning update: made 'in-person' consistent in all files

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse) # a collection of data-related packages
library(knitr) # for kable
library(kableExtra)
library(here)
library(ggbeeswarm)

#reading cleaned data
rhode = read_csv(
  file = here("outputs/data/cleaned_learningmodel/rhode.csv"),       
  show_col_types = FALSE
)

colorado = read_csv(
  file = here("outputs/data/cleaned_learningmodel/colorado.csv"),       
  show_col_types = FALSE
)

connecticut = read_csv(
  file = here("outputs/data/cleaned_learningmodel/connecticut.csv"),       
  show_col_types = FALSE
)

massachusets = read_csv(
  file = here("outputs/data/cleaned_learningmodel/massachusets.csv"),       
  show_col_types = FALSE
)
minnesota = read_csv(
  file = here("outputs/data/cleaned_learningmodel/minnesota.csv"),       
  show_col_types = FALSE
)

mississippi = read_csv(
  file = here("outputs/data/cleaned_learningmodel/mississippi.csv"),       
  show_col_types = FALSE
)

ohio = read_csv(
  file = here("outputs/data/cleaned_learningmodel/ohio.csv"),       
  show_col_types = FALSE
)
virginia = read_csv(
  file = here("outputs/data/cleaned_learningmodel/virginia.csv"),       
  show_col_types = FALSE
)
westvirginia = read_csv(
  file = here("outputs/data/cleaned_learningmodel/westvirginia.csv"),       
  show_col_types = FALSE
)
winscosin = read_csv(
  file = here("outputs/data/cleaned_learningmodel/winscosin.csv"),       
  show_col_types = FALSE
)

wyoming = read_csv(
  file = here("outputs/data/cleaned_learningmodel/wyoming.csv"),       
  show_col_types = FALSE
)

```

# Introduction

The COVID-19 pandemic has profoundly disrupted education systems worldwide, compelling schools and districts across the United States to rapidly adapt their learning models. In 2020 and 2021, educational institutions faced unprecedented challenges, transitioning between in-person, hybrid, and virtual learning models in response to evolving public health guidelines and infection rates. This shift has brought into sharp focus the need to understand the implications of these various schooling models on educational outcomes.

This paper delves into a comprehensive analysis across 11 U.S. states, examining the relationship between different schooling models and their impact on student enrollment and staffing patterns. The unique dataset compiled for this study encompasses a variety of metrics, including total enrollment, in-person and virtual attendance, as well as staff count across different learning models. By analyzing data from kindergarten to 12th grade, this study provides a granular view of how the pandemic has reshaped the educational landscape across these states.

Our findings reveal significant variations in how states have navigated the challenges posed by the pandemic, with notable differences in the adoption of in-person, hybrid, and virtual learning models. We observe that these variations are not only reflective of public health policies but also indicate broader socio-economic and demographic influences. For instance, preliminary analysis suggests that shifts to virtual learning correlate with changes in student enrollment, raising questions about equity and access in education during these challenging times.

The remainder of the paper is structured as follows: Section 2 provides a detailed overview of the data and methodology employed in this study. Section 3 presents an in-depth analysis of the schooling models across the 11 states, offering insights into the trends and patterns observed in the data. Section 4 discusses the key findings, exploring the implications of these schooling models on educational outcomes. Finally, Section 5 concludes with a reflection on the study's findings, limitations, and potential directions for future research in this critical area of educational policy.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

library(dplyr)

# Assuming each state data frame has the same structure and columns
all_states <- bind_rows(
  colorado %>% mutate(state = "Colarado"),
  connecticut %>% mutate(state = "Connecticut"),
  massachusets %>% mutate(state = "Massachusets"),
  minnesota %>% mutate(state = "Minnesota"),
  mississippi %>% mutate(state = "Mississsippi"),
  ohio %>% mutate(state = "Ohio"),
  rhode %>% mutate(state = "Rhode Islands"),
  virginia %>% mutate(state = "Virginia"),
  westvirginia %>% mutate(state = "West Virginia"),
  winscosin %>% mutate(state = "Winscosin"),
  wyoming %>% mutate(state = "Wyoming"),

)

all_states$time_period_start <- as.Date(all_states$time_period_start, format = "%Y-%m-%d")
all_states$learning_model <- as.factor(all_states$learning_model)

library(ggplot2)

ggplot(all_states, aes(x = time_period_start, y = learning_model, color = state)) +
  geom_line() +
  labs(title = "Comparison of Learning Models in US States Over Time",
       x = "Time Period Start", y = "Learning Model") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

\newpage

# References